---
title: "Aides Drug"
author: "Jaelynn Schultz"
date: "Summer 2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
females <- filter(triUmp, gender == "F")
```

```{r eval=FALSE}
females <- females %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H0$ There is no difference between the effects of the drug or the placebo in preventing women from getting aids.

$HA$ There is a difference between the effects of the drug in comparison to the placebo in preventing women from getting aids.  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

In this case, we need to separate the population to showcase the females. Then we need to run the inference against the subset of only females. From this point, we can test the effectiveness of triUmpsal among females. 

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->



## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=females)
barchartGC(~infected + treatment,data=females, type="percent")

```

Viewing the two bar charts above, we can see the frequency of females who got aids is somewhat smaller then those who did not. But after looking at the percentage graph, we can also see that of those females who got aids, most had taken the drug. Basically, the women who took the drug were just as likely to get aids. This suggests that the drug is ineffective against aids in females. Further testing can be done to test this.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=females)
rowPerc(table1)
colPerc(table1)
```

The tables above feature the specific values for the percentages of females who did and did not get the treatment and those who did and did not get aids. Just like the previous information in the graphical descriptive results, this information also shows that majority of the women, whether they took the drug or not, did not get aids. 96.31% of those who took the drug did not get infected (97.58% of the females took the drug). This supports the notion that the drug is ineffective in protecting women against aids. 

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

According to the chi squared test shown above, the p-value is quite small, which means we reject the null hypothesis that claims there is no difference between taking the drug or taking the placebo in how it effects females getting or not getting aids. However, this is the opposite of the expected effect of the drug. Instead of preventing aids, the drugs appears to heighten ones chances of getting it. As such, we can confidently say this drug is not effective in aids prevention in women.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The goal of this research was to analyze whether the drug triUmp is effective in preventing aids infections in women. After comparing this drug to a placebo and reviewing the graphical, numerical, and inferential results, it was determined that the drug actually produces the opposite of the desired effect. Majority of the women who got aids took the drug. In conclusion, the drug proved to be ineffective in preventing women from getting aids. 
